How to \emph{not} write Virtual Machines for Dynamic Languages

### 2007-07-24 talk/dyla2007/talk.tex
something about OO VMs

\begin{frame}
  \frametitle{The Problems of C}
  When writing a VM in C it is hard to reconcile:
  \begin{itemize}
  \item
    flexibility, maintainability
  \item
    simplicity of the VM
  \item
    performance (needs dynamic compilation techniques)
  \end{itemize}
  \pause
  \begin{block}{
    Python Case}
    \begin{itemize}
    \item
      \alert{CPython} is a very simple bytecode VM, performance not great
    \item
      \alert{Psyco} is a just-in-time-specializer, very complex, hard to
      maintain, but good performance
    \item
      \alert{Stackless} is a fork of CPython adding microthreads. It was never
      incorporated into CPython for complexity reasons
    \end{itemize}
  \end{block}
\end{frame}

@@

\begin{frame}
  \frametitle{Title}
  \begin{itemize}
  \item
  \end{itemize}
\end{frame}

>>>>

\begin{frame}
  \frametitle{Implementing VMs in C}
  When writing a VM in C it is hard to reconcile:
  \begin{itemize}
  \item
    flexibility, maintainability
  \item
    simplicity of the VM
  \item
    performance (needs dynamic compilation techniques)
  \end{itemize}
  \pause
  \begin{block}{
    Python Case}
    \begin{itemize}
    \item
      \alert{CPython} is a very simple bytecode VM, performance not great
    \item
      \alert{Psyco} is a just-in-time-specializer, very complex, hard to
      maintain, but good performance
    \item
      \alert{Stackless} is a fork of CPython adding microthreads. It was never
      incorporated into CPython for complexity reasons
    \end{itemize}
  \end{block}
\end{frame}

@@

\begin{frame}
  \frametitle{Implementing Languages on Top of OO VMs}
  \begin{itemize}
  \item
    users wish to have easy interoperation with the OO VMs used by the industry
    (JVM, CLR)
  \item
    bridges don't work that well (XXX maybe skip this)
  \item
    therefore re-implementations of the language on the OO VMs are started
  \item
    even more implementation proliferation
  \item
    implementing on top of an OO VM has its own set of problems
  \end{itemize}
  \pause
  \begin{block}{
    Python Case}
    \begin{itemize}
    \item
      \alert{Jython} is a Python-to-Java-bytecode compiler
    \item
      \alert{IronPython} is a Python-to-CLR-bytecode compiler
    \item
      both are slightly incompatible with the newest CPython version (especially
      Jython)
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Benefits of implementing on top of OO VMs}
  \begin{itemize}
  \item
    higher level of implementation
  \item
    the VM supplies a GC and mostly a JIT
  \item
    better interoperability than what the C level provides
  \item
    some proponents believe that eventually one single VM should be enough
  \end{itemize}
  \pause
  \begin{block}{
    Python Case}
    \begin{itemize}
    \item
      both Jython and IronPython allow easy access to the underlying VMs libraries
    \item
      XXX what else?
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{The problems of OO VMs}
  \begin{itemize}
  \item
    some of the benefits of OO VMs don't work out in practice
  \item
    most immediate problem: it be hard to map concepts of the dynamic lang to
    the host OO VM
  \item
    performance is often not improved, and can be very bad, because of the
    semantic mismatch between the dynamic language and the host VM
  \item
    poor interoperability with everything outside the OO VM
  \item
    in practice, one OO VM is not enough
  \end{itemize}
  \pause
  \begin{block}{
    Python Case}
    \begin{itemize}
    \item
      Jython about 5 times slower than CPython
    \item
      IronPython is about as fast as CPython (but some introspection features
      missing)
    \item
      XXX something about semantic mismatch?
    \end{itemize}
  \end{block}

\end{frame}

\begin{frame}
  \frametitle{}
  \begin{itemize}
  \item
  \end{itemize}
  \pause
  \begin{block}{
    Python Case}
    \begin{itemize}
    \item
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{}
  \begin{itemize}
  \item
  \end{itemize}
\end{frame}
